Methods for Analyzing Models of\\ Rod-Shaped Bacteria

### 2025-11-27 paper/main.tex
move heterogeneity into separate subsection at end of methods

In order to properly estimate the parameters of our model on the single-cell level, we need to be
able to take into account any heterogeneity that may be present.
Therefore, the developed methods are independent of how we assign parameters within our model.
This means we can opt to incorporate cell-cell variability by specifying parameters on a
per-cell basis.
Or we can employ shared parameters with identical numerical values across all agents for properties
which are expected to be identical across the population.
Furthermore, mixed approaches where some parameters are assigned globally and others are chosen
individually can be realized as well.

For our particular model presented in section~\ref{sec:computational-model}, a list of parameters
is presented in table~\ref{table:list-of-parameters} but the following calculations are equally
true for any single-cell model.
We assume that each cell contains $N_p$ number of parameters that need to be assigned numerical
values.
We identify a subset $N_g\leq N_p$ of parameters which will be treated globally, thus with equal
numerical value for every cell.
This means that $N_p - N_g$ parameters will be assigned individually for each of number of cells.
The total amount of parameters $N_\text{tot}$ in the system is then given by
\begin{equation}
    N_\text{tot} = N_g + N_c (N_p - N_g).
\end{equation}
where $N_c$ is the number of unique cells which we encounter over the time interval of interest.
It is clear, that we should avoid, to needlessly treat parameters non-globally since this increases
the size of the parameter space considerably.
The volume of parameter space which needs to be sampled by optimization schemes scales exponentially
with the dimension ie. the total number of parameters $N_\text{tot}$.
Notice that the presented approach reflects our current work but can easily be adapted to other
scenarios where different strategies of parameters assignment are desired.
This can be usefull e.g. when considering two distinct species where the value of a particular
parameter only depends on the species and is thus neither identical across all cells, nor
individual per cell.

@@

\section{Applications}

@@

%###################################################################################################
\subsection{Application 1: Elastic deformations of \acl{ecoli}}

>>>>

@@

% --------------------------------------------------------------------------------------------------
\subsection{The challenge of Cellular Heterogeneity}
\label{sec:role-of-heterogeneity}
In order to properly estimate the parameters of our model on the single-cell level, we need to be
able to take into account any heterogeneity that may be present.
Therefore, the developed methods are independent of how we assign parameters within our model.
This means we can opt to incorporate cell-cell variability by specifying parameters on a
per-cell basis.
Or we can employ shared parameters with identical numerical values across all agents for properties
which are expected to be identical across the population.
Furthermore, mixed approaches where some parameters are assigned globally and others are chosen
individually can be realized as well.

For our particular model presented in section~\ref{sec:computational-model}, a list of parameters
is presented in table~\ref{table:list-of-parameters} but the following calculations are equally
true for any single-cell model.
We assume that each cell contains $N_p$ number of parameters that need to be assigned numerical
values.
We identify a subset $N_g\leq N_p$ of parameters which will be treated globally, thus with equal
numerical value for every cell.
This means that $N_p - N_g$ parameters will be assigned individually for each of number of cells.
The total amount of parameters $N_\text{tot}$ in the system is then given by
\begin{equation}
    N_\text{tot} = N_g + N_c (N_p - N_g).
\end{equation}
where $N_c$ is the number of unique cells which we encounter over the time interval of interest.
This value includes cells which have divided or died during the inspected time interval.

It is clear, that we should avoid, to needlessly treat parameters non-globally since this increases
the size of the parameter space considerably.
The volume of parameter space which needs to be sampled by optimization schemes scales exponentially
with the dimension ie. the total number of parameters $N_\text{tot}$.
Notice that the presented approach reflects our current work but can be adapted easily to other
scenarios where different strategies of parameter assignment are desired.
This can be usefull e.g. when considering two distinct species where the value of a particular
parameter only depends on the species and is thus neither identical across all cells, nor
individual per cell.

%###################################################################################################
\section{Applications}

@@

% --------------------------------------------------------------------------------------------------
\subsection{Application 1: Elastic deformations of \acl{ecoli}}
